---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


```{r setup, echo = FALSE, include=FALSE}
library(Capstone)
library(tidyverse)

```


# Introduction

This package is the result of the Capstone Project of the JHU's Coursera specialization. It contains a few functions that process, clean and plot in an ideal way the information contained in a dataframe with the records of earthquakes registered in our planet, categorized by country, number of deaths, magnitude, etc.

## Data

The mentioned dataframe is included in the package. 

# Functions

## Cleaning

### eq_clean_data

This function takes the entire dataframe as argument, and it creates the variable DATE as a class Date, out of the original variables: YEAR, MONTH and DAY. It also transform the columns LATITUD and LONGITUD to numeric class. This will be useful when we attempt to plot the information.

### eq_location_clean

This function takes the original LOCATION_NAME variable (which contains the country, city and other relevant location information), and eliminates the country information (which is redundant with another original variable called COUNTRY), and transform the strings to title cases (instead of all upcases)

## Plotting: Maps

The following functions also are for plotting information, but instead of timelines we use a World Map. This way, we can visualize each earthquake in the exact location it occurred.

### eq_map

This function takes the earthquake dataframe as input (or a subset of the dataframe), and plots a map with its information, indicating by points the location of each specific earthquake.

### eq_create_label

Finally, this function creates a label to be displayed along with the previous map. This label is a pop-up that appears when you place the mouse on any earthquake point on the map.
