---
title: "MSDR-Capstone"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MSDR-Capstone}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Capstone)
library(tidyverse)

```


# Introduction

This package is the result of the Capstone Project of the JHU's Coursera specialization. It contains a few functions that process, clean and plot in an ideal way the information contained in a dataframe with the records of earthquakes registered in our planet, categorized by country, number of deaths, magnitude, etc.

## Data

The mentioned dataframe is included in the package. To call it we need to use the following commands:


```{r, echo = TRUE}
data("earthquakes_raw")
```

# Functions

## Cleaning

The firs pair of functions are:
* eq_clean_data
* eq_location_clean


### eq_clean_data

This function takes the entire dataframe as argument, and it creates the variable DATE as a class Date, out of the original variables: YEAR, MONTH and DAY. It also transform the columns LATITUD and LONGITUD to numeric class. This will be useful when we attempt to plot the information.

```{r, echo = TRUE}
earthquakes_raw %>% 
  eq_clean_data()
```

### eq_location_clean

This function takes the original LOCATION_NAME variable (which contains the country, city and other relevant location information), and eliminates the country information (which is redundant with another original variable called COUNTRY), and transform the strings to title cases (instead of all upcases)


```{r, echo = TRUE}
earthquakes_raw %>% 
  eq_clean_data() %>% 
  eq_location_clean() %>% 
  select(LOCATION_NAME) %>% 
  drop_na()
```

## Plotting (Geoms): Timelines

The following functions are of the geom class, which are made to be used as the classical ggplo2 geoms, but with new propoerties that are better suited to the type of information we are dealing with.

### geom_timeline

This geom produces a timeline with the earthquakes provided, ordered by country, and also indicating the Richter Scale (EQ_PRIMARY) and the number of total deaths caused by the earthquake (DEATHS).

```{r, echo = TRUE, fig.height = 3, fig.width = 7, fig.align = "center"}
earthquakes_raw %>% 
  eq_clean_data() %>% 
  eq_location_clean() %>%
  select(DATE, EQ_MAG_ML, DEATHS, COUNTRY, LOCATION_NAME, EQ_PRIMARY) %>%
  filter(COUNTRY %in% c("THAILAND", "JAPAN", "INDONESIA"),
         DATE > "2000-01-01") %>% 
  ggplot() +
  geom_timeline(ggplot2::aes(x     = DATE,
                             y     = COUNTRY,
                             size  = EQ_PRIMARY,
                             color = DEATHS))
```

### geom_timeline_label

This geom adds the respective label indicating the location (LOCATION_NAME) to a max number of earthquakes by country (filtered by its magnitude, plotting the location of the n earthquakes with the highest magnitude).

```{r, echo = TRUE, fig.height = 3, fig.width = 7, fig.align = "center"}
earthquakes_raw %>% 
  eq_clean_data() %>% 
  eq_location_clean() %>%
  select(DATE, EQ_MAG_ML, DEATHS, COUNTRY, LOCATION_NAME, EQ_PRIMARY) %>%
  filter(COUNTRY %in% c("THAILAND", "JAPAN", "INDONESIA"),
         DATE > "2000-01-01") %>% 
  ggplot() +
  geom_timeline(ggplot2::aes(x     = DATE,
                             y     = COUNTRY,
                             size  = EQ_PRIMARY,
                             color = DEATHS)) +
  geom_timeline_label(aes(x = DATE, label = LOCATION_NAME, y = COUNTRY, mag = EQ_PRIMARY, n_max = 5))

```

## Plotting: Maps

The following functions also are for plotting information, but instead of timelines we use a World Map. This way, we can visualize each earthquake in the exact location it occurred.

### eq_map

This function takes the earthquake dataframe as input (or a subset of the dataframe), and plots a map with its information, indicating by points the location of each specific earthquake.

```{r, echo = TRUE, fig.height = 3, fig.width = 7, fig.align = "center"}
earthquakes_raw %>%
  eq_clean_data() %>%
  eq_location_clean() %>%
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>%
  eq_map()
```

### eq_create_label

Finally, this function creates a label to be displayed along with the previous map. This label is a pop-up that appears when you place the mouse on any earthquake point on the map.

```{r, echo = TRUE, fig.height = 3, fig.width = 7, fig.align = "center"}
earthquakes_raw %>%
  eq_clean_data() %>%
  eq_location_clean() %>%
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>%
  mutate(popup_text = eq_create_label(.)) %>%
  eq_map(annot_col = 'popup_text')
```


